---
title: "bs823HW4"
author: "Jackson Dial"
date: '2022-10-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

## A

$$
Z = \sqrt{n}(\overline{X} - \mu)/\sigma \sim N(0,1)
$$

$$
\text{since the distribution is symmetric: } z_{\alpha/2} = -z_{1-\alpha/2}
$$
$$
1- \alpha = P[-z_{1-\alpha/2}< \sqrt{n}(\overline{X} - \mu)/\sigma<z_{1-\alpha/2}]
$$

$$
= P[\overline{X} - z_{1-\alpha /2} \space \sigma/\sqrt{n} < \mu<\overline{X} + z_{1-\alpha/2} \space \sigma/\sqrt{n}]
$$

$$
\text{It follows that a 100(1-} \alpha) \% \text{ CI for } \mu \text{ is given by: }
$$

$$
(\bar{x} - z_{1-\alpha / 2} \space \sigma/\sqrt{n}, \bar{x} + z_{1-\alpha/2} \space \sigma \sqrt{n})
$$

## B

```{r}
two.sided.ci <- function(xbar, sigma, alpha, n){
  zscore <- qnorm(1-alpha/2)
  margin_of_error <- zscore * sigma / sqrt(n)
  ub <- xbar + margin_of_error
  lb <- xbar - margin_of_error
  ci <- c(lb, ub)
  return (ci)
}
```

```{r}
two.sided.ci(xbar = 3, sigma = 1.2, alpha = .05, n = 100)
```


## C

To create a simulation study for empirical assessment of coverage probability using $N$ simulation replicates for the confidence interval estimator, I would sample many times from a normal distribution for the $\bar{x}$ value. I would calculate a CI for each of the sample means, and then check to see if the true value of the mean of my distribution was contained in the interval. I would calculate the percentage of confidence intervals that contain the true mean, and this would be my coverage probability.

## D

```{r}

samps <- rnorm(25, 1, .25)
for (samp in samps){
  two.sided.ci()
}

```



## E

## F

## G

